---
title: "Problem2"
runtime: shiny
output: html_document
---

# The Lagrange Polynomial

First we define a function that is going to be taking the coordinates given by the user:
```{r}
library(polynom)
library(pracma)
lagrange <- function(coordinates) {
```

Plot the Lagrange polynomial, evaluated within the x coordinates (from parameter coordinates)

# Parameters
  
coordinates : nx2 matrix 

, where the first column represents the x coordinate while the second column represents the y coordinate.
  
# Returns

The Lagrange polynomial

```{r}
  x = coordinates[,1]
  y = coordinates[,2]
  
  interPoly = poly.calc(x,y)
  xx = x
  yy = lagrangeInterp(x,y,xx)
  
  plot(xx, yy, xlab="x", ylab="f(x)", main = ' Lagrange polynomials ')
  lines(interPoly, col = "#4caf50")
  
  legend(
    'topleft',
    inset = .05,
    legend = c("Coordinates", "Lagrange polynomial"),
    col = c('black', '#4caf50'),
    lwd = c(1),
    bty = 'n',
    cex = .75
  )
  
  return(interPoly)
}
```

Calling the function:

```{r}

x = seq(-5,5,0.5)
y = x^3 + 5*x^2 + 1
lagrange(cbind(x,y))
```



