\documentclass{article}

\usepackage[margin=2cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

\usepackage{mmacells}


\mmaDefineMathReplacement[â‰¤]{<=}{\leq}
\mmaDefineMathReplacement[â‰¥]{>=}{\geq}
\mmaDefineMathReplacement[â‰ ]{!=}{\neq}
\mmaDefineMathReplacement[â†’]{->}{\to}[2]
\mmaDefineMathReplacement[â§´]{:>}{:\hspace{-.2em}\to}[2]
\mmaDefineMathReplacement{âˆ‰}{\notin}
\mmaDefineMathReplacement{âˆž}{\infty}
\mmaDefineMathReplacement{ð••}{\mathbbm{d}}


\mmaSet{
  morefv={gobble=2},
  linklocaluri=mma/symbol/definition:#1,
  morecellgraphics={yoffset=1.9ex}
}


\begin{document}

Input from question with output from my Mathematica version.
Input is in input form, it can be copied and pasted to Mathematica.
\begin{mmaCell}[functionlocal=y]{Code}
  Integrate[{y^(-3)*(1-(a/y)^2)^(-2)},{y,r,Infinity}]
\end{mmaCell}
\begin{mmaCell}{Output}
  \{ConditionalExpression[-\mmaFrac{1}{2 (\mmaSup{a}{2} - \mmaSup{r}{2})},
     Im[r] Re[a] â‰  Im[a] Re[r] || ((a + r > 0 || a + r âˆ‰ Reals) &&
       ((Re[a] < r && Im[a] == 0) || a - r âˆ‰ Reals)) || r âˆ‰ Reals]\}
\end{mmaCell}

For comparison, same cells obtained by including PDFs exported from Mathematica:
% You need inCell.pdf and outCell.pdf files for this to work.
% \mmaCellGraphics{Input}{inCell}
% \mmaCellGraphics[yoffset=3.5ex]{Output}{outCell}

Same input expression, but in standard form (as if it was inputted using math assistant). Note that syntax coloring still works.
\begin{mmaCell}[index=3,functionlocal=y]{Input}
  \mmaSubSupM{\int}{r}{âˆž}\{\mmaFrac{1}{\mmaSup{y}{3}\mmaSup{\big(1-\mmaSup{\big(\mmaFrac{a}{y}\big)}{2}\big)}{2}}\}ð••y
\end{mmaCell}


More features:
\begin{mmaCell}[
  moredefined=f,
  functionlocal=a,
  local=b,
  pattern={x_,x},
  excessargument=n,
  linkbuiltin=List
]{Code}
  (* A (* nested *) comment. *)
  Block[{a=3},a+2]
  f[x_]:=2x+1 
  Module[{b=c},
      Print["a string \" with double quotes inside ", b/d];
      b+1
  ]
  f[z]//FullForm
  Sin[m,n]
  List[1,List[2,3]]; (* Links to documentation. *)
\end{mmaCell}
\begin{mmaCell}{Output}
  5
\end{mmaCell}
\begin{mmaCell}{Print}
  a string " with double quotes inside \mmaFrac{c}{d}
\end{mmaCell}
\begin{mmaCell}[addtoindex=1]{Output}
  1+c
\end{mmaCell}
\begin{mmaCell}[form=FullForm]{Output}
  Plus[1,Times[2,z]]
\end{mmaCell}
\begin{mmaCell}[messagelink={message/General/argx}]{Message}
  Sin::argx: Sin called with 2 arguments; 1 argument is expected. >>
\end{mmaCell}
\begin{mmaCell}{Output}
  Sin[m,n]
\end{mmaCell}
\begin{mmaCell}{Output}
  \{1,\{2,3\}\}
\end{mmaCell}
Graphics mixed with code:
\begin{mmaCell}{Code}
  a Graphics[{Green,Disk[]},ImageSize->50]/2
\end{mmaCell}
% You need greenDisk.pdf file for this to work.
% \begin{mmaCell}[moregraphics={moreig={scale=.7}}]{Output}
%   \mmaFrac{a \mmaGraphics{greenDisk}}{2}
% \end{mmaCell}
\begin{mmaCell}{Code}
  (* Different roles of one symbol. *)
  x;
  _x;
  \mmaDef{x_};
  \mmaPat{x_}->2x;
  \mmaPat{x_}:>2\mmaPat{x};
  f[\mmaPat{x_}]=2x;
  f[\mmaPat{x_}]:=2\mmaPat{x};
  Block[{\mmaFnc{x}},2\mmaFnc{x}];
  Module[{\mmaLoc{x}},2\mmaLoc{x}];
  \mmaUnd{Module}; (* undefined Module symbol *)
  \mmaLnB{Module}; (* Module symbol with link to documentation *)
  Sin[x\mmaExc{,x}];
\end{mmaCell}
\begin{mmaCell}[addtoindex=11]{Input}
  (* Nesting of "formatting boxes" *)
  \mmaSup{a}{\mmaSup{a}{a}} \mmaSub{a}{\mmaSub{a}{a}} \mmaSubSup{a}{\mmaSubSup{a}{a}{a}}{\mmaSubSup{a}{a}{a}} \mmaUnder{a}{\mmaUnder{a}{a}} \mmaOver{a}{\mmaOver{a}{a}} \mmaUnderOver{a}{\mmaUnderOver{a}{a}{a}}{\mmaUnderOver{a}{a}{a}} \mmaFrac{a}{\mmaFrac{a}{\mmaFrac{a}{a}}} \mmaSqrt{\mmaSqrt{a}} \mmaRadical{a}{\mmaRadical{a}{a}}
\end{mmaCell}
\begin{mmaCell}{Input}
  (* Replacements for infix operators in "Input" cells. *)
  x>=y; x<=y; x!=y; x->y; x:>y;
\end{mmaCell}
\begin{mmaCell}{Code}
  (* No replacement in "Code" cells. *)
  x>=y; x<=y; x!=y; x->y; x:>y;
\end{mmaCell}
\begin{mmaCell}[label={(\mmaCellIndex)custom}]{Code}
  (* Cell with custom label. *)
\end{mmaCell}
\begin{mmaCell}{Input}
  \mmaLnT{x}=2;(* labeled definition of x *)
\end{mmaCell}
\begin{mmaCell}{Input}
  \mmaLnL{x}(* usage of x with link to its definition *)
\end{mmaCell}

Implemented syntax elements:
\begin{mmaCell}[
  defined=defined,
  undefined=undefined,
  functionlocal=functionlocal,
  local=local,
  pattern=pattern,
  localconflict=localconflict,
  globalconflict=globalconflict,
  excessargument=excessargument,
  unknownoption=unknownoption,
  unwantedassignment=unwantedassignment,
  shadowing=shadowing,
  syntaxerror=syntaxerror,
  emphasizedsyntaxerror=emphasizedsyntaxerror,
  formattingerror=formattingerror,
]{Code}
  \mmaDef{Def} defined (*  defined symbol *)
  \mmaUnd{Und} undefined (* UndefinedSymbol *)
  \mmaFnc{Fnc} functionlocal (* FunctionLocalVariable *)
  \mmaLoc{Loc} local (* LocalVariable *)
  \mmaPat{Pat} pattern (* PatternVariable *)
  \mmaLCn{LCn} localconflict (* LocalScopeConflict *)
  \mmaGCn{GCn} globalconflict (* GlobalToLocalScopeConflict *)
  \mmaExc{Exc} excessargument (* ExcessArgument *)
  \mmaOpt{Opt} unknownoption (* UnknownOption *)
  \mmaAsg{Asg} unwantedassignment (* UnwantedAssignment *)
  \mmaShd{Shd} shadowing (* SymbolShadowing *)
  \mmaSnt{Snt} syntaxerror (* SyntaxError *)
  \mmaEmp{Emp} emphasizedsyntaxerror (* EmphasizedSyntaxError *)
  \mmaFmt{Fmt} formattingerror (* FormattingError *)
\end{mmaCell}

Inline cell: \mmaInlineCell[functionlocal=a]{Code}{Module[{a=5}, a]}.
Formatted inline cell:
\mmaInlineCell[pattern={x_,x}]{Input}{f[x_]:=\mmaFrac{\mmaSup{x}{2}}{5}}
\subsubsection*{
  Inline cell inside macro argument:
  \mmaInlineCellNonVerb[functionlocal=x]{Code}{Solve[\mmaSqrt{x}==y,x]}
}

\end{document}