---
title: "GSE1739 Dataset"
author: "Antonio Osamu Katagiri Tanaka - A01212611@itesm.mx"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

---

### Dataset details

**Title:**	Severe acute respiratory syndrome expression profile

**Summary:**	Expression profiling of peripheral blood mononuclear cells (PBMC) from 10 adult patients with severe acute respiratory syndrome (SARS). Results provide insight into the host immune response to the SARS coronavirus.

**Organism:**	Homo sapiens

**Platform:	GPL201:** [HG-Focus] Affymetrix Human HG-Focus Target Array

**Citation:**	
Reghunathan R, Jayapal M, Hsu LY, Chng HH et al. *Expression profile of immune response genes in patients with Severe Acute Respiratory Syndrome.* BMC Immunol 2005 Jan 18;6:2. PMID: [15655079](https://www.ncbi.nlm.nih.gov/pubmed/15655079)

**Reference Series:**	GSE1739

**Sample count:**	14

**Value type:**	count

**Series published:**	2005/01/18

**Dataset taken from:** [GDSbrowser](https://www.ncbi.nlm.nih.gov/sites/GDSbrowser), [GSE1739 query](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE1739), and [GSE1739 geo2r](https://www.ncbi.nlm.nih.gov/geo/geo2r/?acc=GSE1739)

---

The GSE1739 was selected for the Bioinformatics class assignments and activities. The reason behind this decision is the similarities that SARS shares with the novel coronavirus COVID-19. The purpose id to work with up-to-date data that is relecant to the current crisis.

---

```{r}
# Clear all objects (from the workspace)
rm(list = ls())

# Suppress Warning messages
options(warn = -1)

# Install packages from Bioconductor as they're not available in CRAN
# if (!requireNamespace("BiocManager", quietly=TRUE))
#     install.packages("BiocManager")
# BiocManager::install("GEOquery")
# BiocManager::install("Biobase")
# BiocManager::install("affy")
```

#### Load the Bioconductor libraries:
```{r}
library(GEOquery)
library(Biobase)
library(limma)
library(affy)
```

#### Command getGEO() and R lists
The command ´rgetGEO()´ is a function from the packages GEOquery (S. Davis and Meltzer 2007) that can download data directly from the [GEO database](http://www.ncbi.nlm.nih.gov/geo/).

The help command ?getGEO provides a description that stating that It directs the download (if no filename is specified) and parsing of a GEO SOFT format file into an R data structure[..]

However, the default command as specified in the Usage section is:

```{r}
# Enable Warning messages
options(warn = 0)
```

